Safe and Robust Formal Synthesis of Digital Controllers for Continuous Plants with Transient Performance Specifications

### 2017-05-28 automatica/automatica.tex
Settling time proof

%\newcommand{\tool}{{\sf DSSynth}\xspace}

@@

{\color{red} Dario: this section must be carefully described: it will be an important contribution}

%-------------------------------
\section{CEGIS of digital control system with transient specifications}
\label{sec:method}
%-------------------------------

>>>>

%\newcommand{\tool}{{\sf DSSynth}\xspace}
\newcommand{\mat}[1]{\boldsymbol{#1}}
\renewcommand{\vec}[1]{\boldsymbol{#1}}
\newtheorem{theorem}{Theorem}
\newtheorem{proof}{Proof}

@@

In a SISO system, there are 2 values that typically define the response-time of the system (using a step response).
\begin{enumerate}
\item The \emph{rise time} is defined as the time it takes the output to reach a percentage (eg $90\%$) of the target value given a step chnage in the input.
\item The \emph{settling time} is defined as the time it takes the output to stabilize within a margin of the target value (eg $10\%$).
\end{enumerate} 
In the case of first and second order systems (with ), the decay is governed by the equations 
\begin{align*}
&y_{decay}=
&\left\{
\begin{array}{cc}
e^{-|\omega| t}& 1^{st} order\\
e^{\zeta |\omega| t}\left(\frac{\sin\left(|\omega| t\sqrt{1-\zeta^2}+\sin^{-1}\sqrt{1-\zeta^2}\right)}{\sqrt{1-\zeta^2}} \right)&\zeta<1\\
e^{-|\omega| t}-|\omega| t e^{-|\omega| t}&\zeta=1\\
\frac{|\omega| e^{-|\omega_2| t}-|\omega_2| e^{-|\omega| t}}{|\omega|-|\omega_2|}&\zeta>1\end{array}\right.
\end{align*}
where $|\omega|$ and $|\omega_2|$ are the magnitudes of the eigenvalues (roots) of the system and  $\zeta=-\cos(\theta)$ where $\theta$ is the polar angle of the conjugate pair if it exists ($\zeta \geq 1$ if it doesn't).
These equations become increasingly complicated for higher order systems. However, we note their strong dependency on the magnitude of the eigenvalues, and use this property to define an upperbound on the decay function.
All of the above equations will decay faster than $e^{-|\omega| t}$ given $|\omega| < |\omega_2|$, hence we may use this function as our upper bound.
\begin{theorem}
Given a set of eigenvalues for a discrete time modle of an $n^{th}$ order system $\lambda_i =e^{\omega_i}: i \in [1\ n]$, the settling time of the system is upperbounded by
\begin{equation}
t_s<-log_{|\overline{\lambda}|}({p_s}) : |\overline{\lambda}| =\max(|\lambda_i|) \wedge |\overline{\lambda}|<1,\ \forall i \in [1\ n]
\label{eq:set_time}
\end{equation}
where $0\leq p_s \leq 1$ is the desired percentage of the target within to remain after $t_s$.
\end{theorem} 
\begin{proof}
Given a closed loop single output LTI system of order $n$ with dynamics
\begin{equation}
\left\{\begin{array}{c}\vec{x}_k=\mat{A}\vec{x}_{k-1}\\ \vec{y}_k=\mat{C}\vec{x}_k\end{array}\right. \Rightarrow \left\{ \begin{array}{c}\vec{x}_k'=\mat{J}\vec{x}_{k-1}'\\ \vec{y}_k=\mat{C}\mat{S}\vec{x}_k'\end{array}\right. : \left\{ \begin{array}{c}\mat{A}=\mat{S}\mat{J}\mat{S}^{-1} \\ \vec{x}_k'=\mat{S}^{-1}\vec{x}_k\end{array}\right.
\end{equation}
the output decay is a linear combination of the decay of eigenvalues or eigenvalue pairs. Hence the overall decay of the system is always smaller than $p=e^{-|\overline{\omega}| t}$ where $0\leq p\leq 1$ is the remaining percentage of the decay expressed. reversing we have $t=-\frac{ln(p)}{|\overline{\omega}|}=-\frac{ln(p)}{ln(|\overline{\lambda}|)}$. replacing for $p_s$ , $t_s$, and $\overline{\lambda}$ we get equation \eqref{eq:set_time}
\qed
\end{proof}
We may now use this limit to express a constraint in terms of the coefficients for our synthesis process.
Let 
\begin{equation}
P=\prod_{i=0}^n (s-\lambda_i)=\sum_{j=0}^n c_js^j.
\end{equation}
We use Jury's criterion to limit $|\lambda_i| \leq 1$. Let us now define a transformation $\lambda_i'=\frac{\lambda_i}{\overline{\lambda}}$. We want to ensure $\lambda_i' \leq 1$ which can be done by applying Jury's criterion to $P'=\prod_{i=0}^n (s-\lambda_i')=\sum_{j=0}^n c_j's^j$ where $c_j'$ are unknown. However, if we replace for $\lambda_i$, we obtain
\begin{equation}
P'=\prod_{i=0}^n (s-\lambda_i')=\prod_{i=0}^n (s-\lambda_i\overline{\lambda})=\sum_{j=0}^n c_j\overline{\lambda}^{n-j}s^j
\end{equation}
where all members are known, we may now synthesise coefficients that ensure the eigenvalues are smaller than a desired magnitue, thus ensuring that the settling time is too.
%-------------------------------
\section{CEGIS of digital control system with transient specifications}
\label{sec:method}
%-------------------------------
